---
title: "PES University, Bangalore"
author:
- "UE20CS312 - Data Analytics - Worksheet 3a - Basic Forecasting Techniques"
- "Designed by Vishruth Veerendranath, Dept. of CSE - vishruth@pesu.pes.edu"
subtitle: Established under Karnataka Act No. 16 of 2013
output:
  word_document: default
  pdf_document: default
urlcolor: blue
editor_options:
  markdown:
    wrap: 72
---
NAME:Adarsh Subhas Nayak 
SRN: PES1UG20CS620
SEC: 'K'
DATE: 30/9/22



### Time Series Data and Basic Forecasting Techniques
Time Series data is any data that is collected at regular time intervals, with changing observations at every time interval. Processing time series data effectively can help gain meaningful insights into how a quantity changes with time. 

Forecasting a quantity into the future is an essential task, that predicts future values at any particular time. Forecasts can be made using various techniques like Exponential Smoothing to much more complex techniques such as Recurrent Neural Networks. Let's try to process time-series data and forecast values using basic techniques!


  
### Task
Let's imagine it is 2012 and you are in the market to buy an Orange Ultrabook Laptop for college. But this laptop is rare to find and expensive. You would want to put your Data Analytics skills to use, and predict the best time to buy your laptop, such that you can get it for the best price! You would also like to predict when the Orange Ultrabook would be in stock and when it would have high demand.

An electronics store collected sales data for their store weekly, from *February 2010* to *October 2012*, a period of 143 months. You have gotten your hands on this, and will use it to predict how the prices will change in the future.


### Data Dictionary

    Date - Date on which data was collected (end of the week)
    Sales - Weekly sales of the store (in $)
    Holiday_Flag - Boolean Flag. 0 for normal week and 1 for holiday season
    Temperature - Average temperature during the week
    Fuel_Price - Average price during the week (in $/gallon)
    CPI - Consumer Price Index
    Unemployment - Average percentage of Unemployment in the city
    Laptop_Demand - Number of Orange Ultrabook laptops sold during the week
    
### Data Ingestion and Preprocessing

- Read the file into a `data.frame` object
```{r csv}

#reading the csv file. 
df <- read.csv('sales.csv')
head(df)
#print(df)
```

- Pick out the `Sales` column in the `data.frame`.
```{r}
sales <- df$Sales
head(sales)
```

- The `ts` function is used to create the `ts` object in R. Frequency is 52 as it is weekly data. 
The start is specified like `start= c(y, m, d)` as we are dealing with weekly data. If it was monthly data we can omit the `d` and for yearly data we can omit the `m` as well.(`c` is the combine function in R)

```{r}
sales_ts <- ts(sales, frequency = 52, start=c(2010, 2, 5))
sales_ts
```
    
- Visualize the Time-Series of `Sales` column
```{r}
plot.ts(sales_ts)
```

### Points
The problems in this worksheet are for a total of 10 points with each problem having a different weightage.



### Problem 1 (1 Point)
Decompose the `Sales` column into trend, seasonal and random components. Plot these components as well (Hint: Look at the `decompose` function).
```{r}
install.packages("forecast")
library(forecast)
#Decomposing a seasonal time series means separating the time series into these three components
tsA<- decompose(sales_ts,"additive")
plot(tsA)

```
```{r}
library(forecast)
tsM<- decompose(sales_ts,"multiplicative")
plot(tsM)
```
#Observation : 

#We choose additive time series over multiplicative time series as the increase in spikes for seasonal component is constant.
#Trend is not calculated for first and last year, Periodicity of seasonal component is found to be one year.
```{r}
seas<- tsA$seasonal
plot(seas)

```
```{r}
tre<-tsA$trend
plot(tre)

```


```{r}
ran<-tsA$random
plot(ran)
```

   
### Problem 2 (3 Points)
- Perform forecasts using Single, Double and Triple Exponential Smoothing. 
- Plot the forecasts of all three forecasts (different colours) against the true values. (Hint: use `lines`) 
- **Use only one function needed for all 3 forecasts**, only changing parameters to get each of the 3 models (Hint: Think about alternate names)
```{r}
#function "ses" gives out the "simple exponential smoothing"
ses<-ses(sales_ts)
#plotting SES
plot(ses)

```
##Simple exponential smoothing does not do well when there is a trend (there will be always bias) , therefore we use double exponential smoothing.
##double exponential smoothing
```{r}
#for double exponential smoothing we use HOlt's linear method.
Hts <- ts(sales, frequency = 12, start=2010)
#the alpha -> smoothness parameter for the level (trend-less series), we have a beta parameter which controls the smoothness of the trend component. 
HoltModel <- holt(Hts, h=6, alpha=0.2, beta=0.1)

summary(HoltModel)
plot(HoltModel)

```

#Inorder to consider seasonal changes as well as trends we use Triple ES.
# triple exponential smoothing
#lets plot for both additive and multiplicative method.
```{r}
#
HWModelA <- hw(Hts, h=12, alpha=0.2, beta=0.1, gamma=0.25, seasonal =
"additive")
summary(HWModelA)
plot(HWModelA)
```
```{r}
HWModelM <- hw(Hts, h=12, alpha=0.2, beta=0.1, gamma=0.25, seasonal =
"multiplicative")
summary(HWModelM)
plot(HWModelM)
```
#Plotting all of them in a single graph.
```{r}
#dev.new(width=400, height=250, unit="px")
plot(ses,col="green")
par(new=TRUE)
plot(HoltModel,col="blue")
par(new=TRUE)
plot(HWModelA, col="orange" )

```


### Problem 3 (2 Points)
- Forecast `Sales` values by Regression using all other columns. Print summary of regression model. 
- Plot the predicted values against original as well. (Hint: Regression model predictions will not be a Time Series, so need to get both to common index/x-axis)
- (Hint: Will not work unless one column is dropped/transformed before including it in the regression. Use the `lm` function to get linear model)

Note: This is Multiple Linear Regression, that is, using all the columns for regression
```{r}
sales_lm <- lm(sales~ Holiday_Flag + Temperature + Fuel_Price + CPI + Unemployment + Laptop_Demand , data=df)
sales_lm%>% summary()
```
```{r}
##obtaining the predicted values.
library(tidyverse)
explan_sales<-tibble(df)
#head(explan_sales)
predict(sales_lm,explan_sales)

```
```{r}
#now let's plot the predicted values against the observed or the original values.
plot(predict(sales_lm), df$sales,
     xlab = "Predicted Values",
     ylab = "Observed Values")
abline(a = 0, b = 1, lwd=2,
       col = "green")
```




### Problem 4 (2 Points)
Plot the `Laptop_Demand` column as a time series. Identify the forecasting required for this type of Time-series, and forecast the values for all 143 weeks (Hint: Look at functions in the `forecast` package)
```{r}
ts1 <- ts(df$Laptop_Demand, frequency = 12, start = 2012)
plot(ts1)

```
```{r}
tsdA1 <- decompose(ts1, "additive")
plot(tsdA1)
```
```{r}
tsdM1 <- decompose(ts1, "multiplicative")
plot(tsdM1)
```
```{r}
#USING CROSTON MODEL 
Cts <- ts(df$Laptop_Demand, frequency = 12, start = 2010)
plot(Cts)

```
```{r}
CrostonModel <- croston(Cts, h=4, alpha=0.2)
summary(CrostonModel)
plot(CrostonModel)
```






### Problem 5 (2 Points)
Evaluate the accuracy of all 3 Exponential Smoothing models (from Problem 2) and Regression models using the MAPE and RMSE metrics. 
Comment on which is the best Exponential Smoothing method, and if Regression is better than Exponential Smoothing? 
Provide a reasoning for why the best model is better suited for the `Sales` data (Bonus Point: reasoning for why the 2 other models perform similarly)

```{r}
#lets take a summary of the "Holt Model".
summary(HoltModel)
```
```{r}
summary(HWModelA)
```
```{r}
summary(ses)
```
##Has least MAPE and RMSE so seems to be the best method
