---
title: |
  PES University, Bangalore
subtitle: "**UE20CS312 - Data Analytics**\n\n**Worksheet 2b : Multiple Linear Regression**\n\nCourse Anchor : Dr. Gowri Srinivasa\n\n\nPrepared by : Nishanth M S - nishanthmsathish.23@gmail.com\n"
output:
  pdf_document:
    fig_width: 6
    fig_height: 6
  html_document:
    df_print: paged
urlcolor: blue
editor_options:
  markdown:
    wrap: 72
---
---
Name : Adarsh S Nayak
SRN : PES1UG20CS620
Roll No : 54
---

## Multiple Linear Regression

Multiple Linear Regression (MLR) is a statistical technique that uses
several explanatory variables to predict the outcome of response
variable.The goal of MLR is to model **a linear relationship** between
explanatory(independent) variables and response(dependent) variables.

## Data Dictionary

The data required for this worksheet can be downloaded [from this GitHub
Link](https://github.com/Data-Analytics-UE20CS312/Unit-2-Worksheets/tree/main/2b%20-%20Multi%20Linear%20Regression).
The data was obtained from
[this](https://www.kaggle.com/datasets/bricevergnou/spotify-recommendation)
dataset from Kaggle. The dataset contains features of songs on Spotify
collected using Spotify API.The features are as follows :

\-**acousticness** : A confidence measure from 0.0 to 1.0 of whether the
track is acoustic. 1.0 represents high confidence the track is acoustic.

\-**danceability** : Danceability describes how suitable a track is for
dancing based on a combination of musical elements including tempo,
rhythm stability, beat strength, and overall regularity. A value of 0.0
is least danceable and 1.0 is most danceable.

\-**duration_ms** : The duration of track in milliseconds.

\-**energy** : Energy is a measure from 0.0 to 1.0 and represents a
perceptual measure of intensity and activity.Perceptual features
contributing to this attribute include dynamic range, perceived
loudness, timbre, onset rate, and general entropy.

\-**instrumentalness** : Predicts whether a track contains no vocals.The
closer the instrumentalness value is to 1.0, the greater likelihood the
track contains no vocal content. Values above 0.5 are intended to
represent instrumental tracks, but confidence is higher as the value
approaches 1.0.

\-**key** : The key the track is in. Integers map to pitches using
standard Pitch Class notation

\-**liveness** : Detects the presence of an audience in the recording.
Higher liveness values represent an increased probability that the track
was performed live. A value above 0.8 provides strong likelihood that
the track is live.

\-**loudness** : The overall loudness of a track in decibels (dB).
Loudness values are averaged across the entire track and are useful for
comparing relative loudness of tracks. Loudness is the quality of a
sound that is the primary psychological correlate of physical strength
(amplitude). Values typical range between -60 and 0 db.

\-**mode** : Mode indicates the modality (major or minor) of a track,
the type of scale from which its melodic content is derived. Major is
represented by 1 and minor is 0.

\-**speechiness** : Speechiness detects the presence of spoken words in
a track. The more exclusively speech-like the recording (e.g. talk show,
audio book, poetry), the closer to 1.0 the attribute value. Values above
0.66 describe tracks that are probably made entirely of spoken words.
Values between 0.33 and 0.66 describe tracks that may contain both music
and speech, either in sections or layered, including such cases as rap
music. Values below 0.33 most likely represent music and other
non-speech-like tracks.

\-**tempo** : The overall estimated tempo of a track in beats per minute
(BPM). In musical terminology, tempo is the speed or pace of a given
piece and derives directly from the average beat duration.

\-**time_signature** : An estimated overall time signature of a track.
The time signature (meter) is a notational convention to specify how
many beats are in each bar (or measure).

\-**valence** : A measure from 0.0 to 1.0 describing the musical
positiveness conveyed by a track. Tracks with high valence sound more
positive (e.g. happy, cheerful, euphoric), while tracks with low valence
sound more negative (e.g. sad, depressed, angry).

Throughout the course of this worksheet , our response variable is
energy. We shall try and apply the concepts learnt in class to predict
the energy of a song using the other features of a song.

## Libraries used

-tidyverse

-corrplot

-olsrr :
[documentation](https://www.rdocumentation.org/packages/olsrr/versions/0.5.3)

## Points
The problems for this worksheet is for a total of 10 points and the weightage is not uniformly distributed.

-   *Problem 1* : 0.5 points
-   *Problem 2* : 2 points
-   *Problem 3* : 2 points
-   *Problem 4* : 1 point
-   *Problem 5* : 1.5 points
-   *Problem 6* : 1 point
-   *Problem 7* : 2 points

Loading packages:
```{r}
library(tidyverse)
library(corrplot)
library(olsrr)
library(ggpubr)
library(dplyr)
library(ggplot2)
library(broom)
library(car)
library(AICcmodavg)
```

## Loading the Dataset

After downloading the dataset and ensuring the working directory is
right , we read the csv into the dataframe.

```{r dataload, message=FALSE, warning=FALSE, results=TRUE }
library(tidyverse)
spotify_df <- read_csv("spotify.csv")
head(spotify_df)
```

## Problem-1    (0.5 Points)

Check for missing values in the dataset and normalize the dataset.

```{r}
#Check for missing values in each column
sapply(spotify_df,anyNA)
#count of missing values in each column
sapply(spotify_df, function(x) sum(is.na(x)))
```

## Problem-2    (2 Points)

Fit a linear model to predict the *energy* rating using *all* other
attributes.Get the summary of the model and explain the results in detail.[*Hint*
: Use the lm() function. [Click
here](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/lm)
To get the documentation of the same.]

```{r}
full_model<-lm(energy~.,data=spotify_df)
summary(full_model)
```
**Summary**:The attributes danceability,loudness,acousticness,instrumentalness and valence are all significant.
That is they are important predictors in determining energy with alpha set to 0.05.
Larger the adjusted R-squared and smaller the residual standards error implies better the model.
R-squared value is always less than adjusted R-squared value.
F-statistic is analysis of whole model,in this case not all beta are zero.



## Problem-3    (2 points)

With the help of a correlogram and scatter plots, choose the features you
think are important and model an MLR. Justify your choice and explain the new findings.
```{r}
correlation<-cor(spotify_df)
corrplot(correlation,method = 'number')
```
Energy is strongly positively correlated to loudness and accoustiness.
And accoustiness and loudness are also negatively correlated.Implies comparing energy and loudness is sufficient to draw insights about accoustiness.

```{r}
plot(x=spotify_df$loudness,y=spotify_df$energy,xlab="loudness",ylab="energy",main="energy vs loudness")
```
ln(energy) vs x
```{r}
plot(x=spotify_df$loudness,y=log(spotify_df$energy),xlab="loudness",ylab="energy",main="energy vs loudness")
plot(x=spotify_df$instrumentalness,y=spotify_df$energy,xlab="instrumentalness",ylab="energy",main="energy vs instrumentalness")
```
bad model

```{r}
plot(x=spotify_df$acousticness,y=spotify_df$energy,xlab="accousticness",ylab="energy",main="energy vs accousticness")
```

```{r}
reduced_model<-lm(energy~loudness+acousticness,data=spotify_df)
summary(reduced_model)
```
Since adjusted Rsquared value is higher implies its a good model.



## Problem-4    (1 Point)

Conduct a partial F-test to determine if the attributes not chosen by
you in *Problem-3* are significant to predict the energy.What are the
null and alternate hypotheses? [ *Hint* : Use the anova function between
models created in *Problem-2* and *Problem-3*]
```{r}
anova(reduced_model,full_model)
```
Rejecting null hypothesis(since p value <0.05).Among the parameters left out,they are significant
## Problem-5    (1.5 Points)

AIC - Akaike Information Criterion is used to compare different models
and determine the best fit for the data. The best-fit model according to
AIC is the one that explains greatest amount of variation using the
fewest number of attributes. Check
[this](https://www.scribbr.com/statistics/akaike-information-criterion/)
resource to learn more about AIC.

Build a model based on AIC using Stepwise AIC regression.Elucidate your observations from the new model. ( *Hint* : Use
an appropriate function in
[olsrr](https://www.rdocumentation.org/packages/olsrr/versions/0.5.3)
package.)
```{r}
full_model<-lm(energy~.,data=spotify_df)
stepwise_model<-lm(energy~loudness+acousticness+danceability+valence+instrumentalness+mode+key,data=spotify_df)
reduced_model<-lm(energy~loudness+acousticness,data=spotify_df)

models<- list(full_model,reduced_model,stepwise_model)
models.names<- c('full_model','reduced_model','stepwise_model')
aictab(cand.set=models,modnames=models.names)

```
**Analysis**:Lower the value of AIC, better is the model.Therfore stepwise model is the best amongst full model and reduced model.


## Problem-6    (1 Point)

Plot the residuals of the models built till now and comment on it
satisfying the assumptions of MLR.
```{r}
stepwise_model<-lm(energy~loudness+acousticness+danceability+valence+instrumentalness+mode+key,data=spotify_df)
summary(stepwise_model)
```
```{r}
print("full model residuals")
plot(full_model$residuals,pch=16,col="red")
abline(h=0,lty=2)
ols_plot_resid_hist(full_model)

```
Histogram appears to be normally distributed.
```{r}
print("reduced residuals plots")
plot(reduced_model$residuals,pch=16,col="red")
abline(h=0,lty=2)
ols_plot_resid_hist(reduced_model)

```
```{r}
print("stepwise residuals plots")
plot(stepwise_model$residuals,pch=16,col="red")
abline(h=0,lty=2)
ols_plot_resid_hist(stepwise_model)

```
**Comment**:Stepwise model is better compared to full model and reduced model.

## Problem-7    (2 Points)
For the model built in **_Problem-2_** , determine the presence of multicollinearity using VIF. Determine if there are outliers in the data using [Cook's Distance](https://www.statisticshowto.com/cooks-distance/). If you find any , remove the outliers and fit the model for _Problem-2_ and see if the fit improves. [ _Hint_ : All the relevant functions can be found in _olsrr_ package. An observation can be termed as an outlier if it has a Cook's distance of more than 4/n where n is the number of records.] 


```{r}
ols_vif_tol(full_model)
cookd<-ols_plot_cooksd_chart(full_model)
```
Nothing is violating threshold(value of threshold being 4/n).

#remove outliers
```{r}
new_df<-spotify_df[-c(30,35,79,84,114,120,127,144,153,159,171,172,187,191),]
new_full_model<-lm(energy~.,data=new_df)
summary(new_full_model)
```








